### 2023-07-12 R/models.qmd

---
title: "Supplement B: Fitting Models"
date: today
format:
  html: 
    embed-resources: true
    fig-align: center
    fig-dpi: 300
    fig-responsive: true
    number-sections: true
    number-depth: 3
    toc: true
    toc-depth: 3
    smooth-scroll: true
    theme: cosmo
    highlight-style: a11y
execute: 
  echo: true
  warning: false
  error: false
---

<!-- 
RStudio/Quarto is having some issues with figure sizing, 
so using this hacky CSS solution. 
-->

```{css}
#| echo: false

.img-45 {
  width: 4.5in;
}

.img-65 {
  width: 6.5in;
}

```

## Overview

__Goal__: model the spatial distribution of ranch clusters and populations.

__Data__: Dr. Macfarlan's ethnographic data, along with ecological and socio-economic variables.

__Method__: binomial GLM for occupied vs abandoned, negative binomial for population

We want to know what motivates market-integrated, subsistence ranchers to live where they do. This is framed as a trade-off between market integration and ecology.

## R Preamble

```{r libraries}

library(MASS)
library(broom)
library(car)
library(dplyr)
library(ggplot2)
library(here)
library(gt)
library(patchwork)
library(purrr)
library(sf)
library(sfdep)
library(tidyr)
library(viridis)

```

Specify consistent plot theme here:

```{r}
#| code-fold: true

theme_set(theme_bw(12))

theme_update(
  panel.grid = element_blank(),
  plot.margin = margin(2,2,2,2),
  strip.background = element_blank(),
  strip.text = element_text(hjust = 0, size = rel(1.1))
)

```

And some defaults for table formatting:

```{r}
#| code-fold: true

rename_models <- function(x){

  x |> 
    mutate(
      model = case_when(
        model == "occupied"   ~ "Occupied-Abandoned",
        model == "population" ~ "Population",
        model == "secure"     ~ "Population [+ secure]",
        model == "securex"    ~ "Population [x secure]",
        TRUE ~ NA
      )
    )

}

format_double <- function(x){

  x |> mutate(across(where(is.double), \(.x){ round(.x, digits = 3) }))

}

model_table <- function(x){

  x |> 
    gt() |> 
    tab_options(
      column_labels.font.weight = "bold",
      data_row.padding = px(6),
      heading.align = "left",
      heading.border.bottom.style = "hidden",
      row.striping.background_color = "gray96",
      row.striping.include_stub = FALSE,
      row.striping.include_table_body = TRUE,
      table.align = "left",
      table.border.top.style = "hidden",
      table.border.bottom.style = "hidden"
    ) |> 
    sub_missing(missing_text = "")

}

```

## Data

### Geopackage

Specify path to geopackage database holding all spatial vector data.

```{r}

gpkg <- here("data", "choyero.gpkg")

```

### Dependent variables

* Occupied vs Abandoned ranches
* Population size of occupied ranches

Rancheros live in "ranch clusters" occupied by close kin.

```{r}

ranches <- read_sf(gpkg, layer = "ranches")

clusters <- read_sf(gpkg, layer = "clusters")

```

:::{.mx-auto .my-4 .img-65}
![](../figures/overview-map.png)
:::

### Independent variables

* Market-integration: cost-distance (sort of) to La Paz and Constitucion (hours)  
* Ecology: cost-distance to springs (hours)

Cost-distance to La Paz/Ciudad Constitucion provides a proxy measure of access to markets where rancheros sell livestock and buy food and equipment. Cost-distance to springs represents habitat suitability for small-scale pastoralist economies. Measuring it is a complicated matter, as property rights are involved and multiple springs are used. To avoid some of that complexity, we take the average travel time from a ranch cluster to its two nearest springs. So, this is our basic trade-off between market-integration and local ecology. For further details about how we estimate this variable, see "R/leastcostpaths.html."

```{r}

path_to_cities <- read_sf(gpkg, layer = "path_to_cities")

path_to_springs <- read_sf(gpkg, layer = "path_to_springs")

security <- ranches |>   
  st_drop_geometry() |>
  select(cluster, land) |> 
  mutate(secure = ifelse(land == "secure", 1, 0)) |> 
  group_by(cluster) |> 
  summarize(secure = sum(secure)/n())

clusters <- 
  clusters |> 
  mutate(
    springs = path_to_springs$cost,
    cities = path_to_cities$cost,
    paz = path_to_cities$time_to_paz,
    secure = security$secure
  )

remove(path_to_cities, path_to_springs, security, ranches)

```

## Analysis

### Models

In addition to a binomial model of the status of each ranch cluster (occupied vs abandoned), we fit a negative binomial model of the population count of occupied ranch clusters. The negative binomial is an alternative to standard Poisson models where there is an expectation of dispersion in the residuals. The Poisson model assumes that the variance is equal to the mean, but this is not always the case.

```{r}

models <- list(

  occupied = glm(
    occupied ~ paz + springs,
    family = binomial,
    data = clusters
  ),
  population = glm.nb(
    population ~ cities + springs,
    data = clusters |> filter(population > 0) 
  ),
  secure = glm.nb(
    population ~ (cities + springs) + secure,
    data = clusters |> filter(population > 0) 
  ),
  securex = glm.nb(
    population ~ (cities + springs) * secure,
    data = clusters |> filter(population > 0)
  )

)

```

### ANOVA (LRT)

We conduct a Likelihood Ratio Test of the population models to see if adding land security makes a significant improvement to the model.

```{r}

lrt <- with(models, anova(population, secure, securex, test = "LRT"))

```

### Residual Autocorrelation

We use Monte Carlo simulations of Moran's I to test for spatial autocorrelation in the untransformed residuals of each model.

```{r}

get_moran <- function(x, .sf){

  is_nb <- grepl("Negative Binomial", x$family[["family"]])

  if (is_nb) .sf <- .sf |> filter(occupied == 1)

  geometry <- st_geometry(.sf)

  neighbors <- st_knn(geometry, k = nrow(.sf)-1)
  weights <- st_inverse_distance(neighbors, geometry)

  moran <- global_moran_perm(residuals(x), neighbors, weights)

  moran |> 
    tidy() |> 
    mutate(variable = "residuals", .before = 1)

}

morans_i <- models |> 
  map(\(.x){ get_moran(.x, clusters) }) |> 
  bind_rows(.id = "model")

```

### Variance Inflation

A model of variance inflation tests for multi-collinearity in the predictors.

```{r}

vif <- models |> 
  map(vif) |> 
  bind_rows(.id = "model")

```

## Results

### Model evaluation

```{r}
#| code-fold: true

model_statistics <- models |> 
  map(glance) |> 
  # glance returns logLik column as numeric for glm and 'logLik" for negbin
  map(\(.x){ mutate(.x, across(everything(), as.numeric)) }) |> 
  bind_rows(.id = "model") |> 
  format_double() |> 
  rename_models() |> 
  rename("n.obs" = nobs) |> 
  relocate(model)

model_statistics |> 
  model_table() |> 
  tab_header(title = "Model Statistics")

```

```{r}
#| code-fold: true

coefficient_estimates <- models |> 
  map(tidy) |> 
  bind_rows(.id = "model") |> 
  format_double() |> 
  rename_models() |> 
  rename("z value" = statistic) |>
  rename_with(gsub, pattern = "\\.", replacement = " ")

coefficient_estimates |> 
  group_by(model) |> 
  model_table() |> 
  tab_options(row_group.as_column = TRUE) |> 
  tab_stubhead("model") |> 
  tab_header(title = "Coefficient Estimates")

```

```{r}
#| code-fold: true

lrt <- lrt |> 
  as_tibble() |> 
  rename_with(tolower) |> 
  format_double() |> 
  mutate(
    model = case_when(
      model == "cities + springs"            ~ "Population",
      model == "(cities + springs) + secure" ~ "Population [+ secure]",
      model == "(cities + springs) * secure" ~ "Population [x secure]",
      TRUE ~ NA
    )
  )

lrt |> 
  model_table() |> 
  tab_header(
    title = "Likelihood Ratio Test",
    subtitle = "Alternative: difference in log-likelihood greater than zero"
  )

```

```{r}
#| code-fold: true

morans_i <- morans_i |> 
  format_double() |> 
  rename_models() |> 
  rename("rank" = parameter)

morans_i |> 
  select(model, statistic, rank, p.value) |> 
  model_table() |> 
  tab_header(
    title = "Monte Carlo Simulations of Moran's I",
    subtitle = "Alternative: spatial autocorrelation (two-sided)"
  )

```

```{r}
#| code-fold: true

vif <- vif |> 
  format_double() |> 
  rename_models()

vif |> 
  model_table() |> 
  tab_header(
    title = "Variance Inflation",
    subtitle = "Alternative: "
  )

```

High VIF scores for interaction terms can be ignored.

```{r}
#| echo: false

save(
  model_statistics, coefficient_estimates, morans_i, vif,
  file = here("data", "model-statistics.rdata")
)

```

### Marginal Response

Here we calculate the marginal response of each response variable relative to change in each focal variable while holding all non-focal variables at their mean.

```{r}

# to make the lines smoother when plotting
densify <- function(x, n = 300) seq(min(x), max(x), length = n)

template <- with(
  clusters, 
  tibble(
    paz = densify(paz),
    cities = densify(cities),
    springs = densify(springs)
  )
)

estimate <- function(model, newdata){

  est <- predict(
    model, 
    newdata = newdata, 
    se.fit = TRUE
  )

  inv_link <- family(model)$linkinv

  est <- with(
    est,
    tibble(
      y  = inv_link(fit),
      hi = inv_link(fit + 2*se.fit),
      lo = inv_link(fit - 2*se.fit)
    )
  )

  bind_cols(newdata, est)

}

responses <- tibble(
    model = c("occupied", "occupied", "population", "population"),
    variable = c("paz", "springs", "cities", "springs")
  ) |> 
  mutate(
    model_object = models[model],
    data = list(template),
    data = map2(data, variable, \(.x, .y){ .x |> mutate(across(-all_of(.y), mean)) }),
    estimate = map2(model_object, data, estimate),
    estimate = map2(estimate, variable, \(.x, .y){ rename(.x, "x" = all_of(.y)) })
  ) |> 
  select(model, variable, estimate) |> 
  unnest(estimate) |> 
  select(model, variable, x, y, hi, lo)

```

```{r}
#| code-fold: true
#| fig-asp: 1.1

# add old data to display points
observations <- map(models, \(.x){ .x$model })

observations <- tibble(
    model = c("occupied", "occupied", "population", "population"),
    variable = c("paz", "springs", "cities", "springs")
  ) |> 
  mutate(
    old_data = observations[model],
    old_data = map2(old_data, variable, \(.x, .y){ rename(.x, "x" = all_of(.y)) }),
    old_data = map2(old_data, model,    \(.x, .y){ rename(.x, "y" = all_of(.y)) })
  ) |> 
  unnest(old_data) |> 
  select(model, variable, x, y)

baja_labels <- c(
    "paz" = "To La Paz",
    "springs" = "To Springs",
    "cities" = "To Cities"
)

gg_occupied <- ggplot() +
  geom_ribbon(
    data = responses |> filter(model == "occupied"),
    aes(x, ymin = lo, ymax = hi),
    fill = "#F5F5F5",
    color = "transparent"
  ) +
  geom_line(
    data = responses |> filter(model == "occupied"),
    aes(x, y),
    linewidth = 0.7
  ) +
  geom_point(
    data = observations |> filter(model == "occupied"),
    aes(x, y),
    size = 2.5,
    color = alpha("#2E4756", 0.7)
  ) +
  labs(
    x = "Hours",
    y = "Abandoned-Occupied"
  ) +
  facet_wrap(
    vars(variable), 
    scales = "free_x",
    labeller = labeller(variable = baja_labels)
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )

gg_population <- ggplot() +
  geom_ribbon(
    data = responses |> filter(model == "population"),
    aes(x, ymin = lo, ymax = hi),
    fill = "#F5F5F5",
    color = "transparent"
  ) +
  geom_line(
    data = responses |> filter(model == "population"),
    aes(x, y),
    linewidth = 0.7
  ) +
  geom_point(
    data = observations |> filter(model == "population"),
    aes(x, y),
    size = 2.5,
    color = alpha("#2E4756", 0.7)
  ) +
  labs(
    x = "Hours",
    y = "Population"
  ) +
  facet_wrap(
    vars(variable),
    scales = "free_x",
    labeller = labeller(variable = baja_labels)
  )

gg <- gg_occupied/gg_population

ggsave(
  plot = gg,
  here("figures", "response-plots.png"),
  dpi = 600,
  width = 5.75,
  height = 5.75
)

```

:::{.mx-auto .my-4 .img-45}
![](../figures/response-plots.png)
:::

## Session Info

```{r}
#| code-fold: true

# save the session info as an object
pkg_sesh <- sessioninfo::session_info(pkgs = "attached")

# inject the quarto info
pkg_sesh$platform$quarto <- paste(
  quarto::quarto_version(), 
  "@", 
  quarto::quarto_path()
  )

# print it out
pkg_sesh

```