### 2024-01-24 manuscript/reviews-response.md
reviews and response to reviews

We would like to thank the two reviewers and the editor for their time and
effort put into reviewing our manuscript. They offered a number of important
comments that we think make the paper better as a whole. We made substantial
revisions to the methods section as requested by reviewer 2. Reviewer 1 was more
concerned with conceptual and theoretical issues in the paper. Responding to
their concerns required some substantial changes to the structure of the paper,
including new sections in the intro, methods, and discussion.

## EDITOR

### MAJOR COMMENTS

1.  The abstract has a little shortfall compared with the richness of the body.
    Please especially enrich the methodology, result, and discussion parts.

    Response: The abstract was substantially revised to add requested detail and
    to reflect changes made in response to reviewer concerns.

2.  The current Introduction includes some component of methodology (GLM and
    model variables) as the reviewer #2 says. Please restructure the story line
    of Introduction by referring some comments of the reviewer #1. And in the
    last paragraph of Introduction please explicitly state the research
    questions, not methodological advantages in the current manuscript.

    Response: Discussion of methodology has been removed from the introduction
    and the primary research question restated at the end of the introduction
    for emphasis. Please see our comments below for how we address reviewer 1's
    concerns.

3.  The current result table1 is too small for readers. Please add the full
    statistical tables in the supplementary B 5.1 section. It is not so annoying
    but gives rich information to the readers. Additionally, It looks that the
    analysis of "land secure" factor disappeared somewhere in the manuscript.
    Please enhance the description of land fragmentation part.

    Response: Model results table was expanded to include all models.

4.  Both of the reviewers recommended to introduce additional references in
    Discussion part. Please recheck that Discussion contains the unconscious
    assumptions or conclusion without evidences.

    Response: Done. See below for more details about references we added here
    and throughout the paper.

### MINOR COMMENTS

1.  Can you show some photo of the landscape of the target regions to help the
    SUST readers understanding if possible?

    Response: Done!

2.  Please add some references of "Djikstra's algorithm" and "Tobler's Law" and
    other concepts without references if contained.

    Response: References added.

## REVIEWER 1

1.  Is the economy really subsistence based?

    Response: We were over-zealous in our description. They do produce a
    substantial portion of their own diet, but they are also not entirely
    dependent on what they produce, so it is probably more appropriate to refer
    to these communities as "smallholder" ranching economies. To address this,
    we have replaced "subsistence" with "smallholder" everywhere in the text. We
    also emphasize that they were never fully isolated from markets.

2.  Are all springs equally attractive to settlement?

    Response: Indeed, springs in this area are effectively the same. Lerback et
    al (2023) provides a fine-grained analysis of spring systems in the project
    area. They are primarily lentic pools that occur in the arroyos, so it's
    ideal to setup your ranch downslope of the spring on a flat piece of land
    slightly above the low point, so their house doesn't flood.

3.  Land fragmentation vs land insecurity.

    Response: The reviewer is right that these are not the same thing, so we
    have tried to make that more clear in the text. Basically, we see both land
    fragmentation and land insecurity as outcomes of the same process of
    privatization that was spawned by the Agrarian Land Reform of 1992 (in
    conjunction with NAFTA). There's a patchwork of property rights regimes out
    there that have consequences for land security/ownership. We define land
    (in)security in terms of whether individuals have explicit governmental
    documentation identifying rights to their land. Before the 1992 Land Reforms
    and NAFTA, land title to a specific piece of land wasn't really an issue
    because the land was treated as one contiguous commons co-managed by all the
    individual families (who were by and large related to one another through
    blood and marriage). As a result, very few individuals obtained
    documentation to land or became ejido members. After the 1992 Land Reforms,
    however, there were monetary incentives for ejido members to split up and
    sell land. This then raised issues for families that either never had clear
    title to land or never became members of the ejido. Once the land became
    parcelized and placed on land markets, individuals and organizations from
    outside the community began purchasing land (and sometimes fencing it off),
    creating an existential threat for land dis-appropriation. As a result, land
    fragmentation and land security are heavily related to one another today.

4.  Move "historical antecedents" to introduction.

    Response: As requested, additional historical context was added to the
    introduction. We also added a site content section to the methods to
    emphasize aspects of the region most relevant to the analysis.

5.  What about road condition and transport efficiency?

    Response: Road condition is incorporated into the cost-distance analysis
    under the assumption that travel is faster over maintained roads. That said,
    there's only really one way in or out of each watershed (north and south),
    so it's really how close they are to the maintained road within the study
    area that makes the biggest difference.

6.  Is the ecological discussion relevant?

    Response: Our goal is to show that these ranchers must navigate a trade-off
    between distant market services and local ecosystem services, and even as
    they integrate more with markets, they are still constrained by local
    ecology, which puts them in a precarious position. So, understanding the
    ecology, how it shapes their decisions and is shaped by their decisions is
    critically important. We have to tried to make the emphasis on this
    trade-off clearer by changing the title of the paper. It is also emphasized
    more in the abstract and in the introduction.

7.  What about intermediate markets?

    Response: As far as we know, there are no intermediate markets between these
    communities and the major population centers of Ciudad Constitucion and La
    Paz.

## REVIEWER 2

1.  Why binomial rather than logit GLM?

    Response: In R, the default link for the binomial GLM is the logit link. We
    should have been more clear about this, so we added detail to the methods to
    clarify.

2.  Model specification and functional form should be provided in the text.

    Response: The reviewer is absolutely right to request this, so we have added
    detail to the methods section to elaborate on this.

3.  Provide a table that defines the variables used along with some of their
    descriptive statistics.

    Response: This is a solid suggestion, so the table has been added.

4.  Land fragmentation vs land insecurity.

    Response: see response to reviewer 1.

5.  Assumptions of travel time analysis need citations.

    Response: The reviewer requested examples of the use of travel time in
    ethnographic and related studies involving market integration. We actually
    provided some of these, but they were not in the methods section on travel
    time estimation, which was probably a mistake on our part, so we have moved
    them there. We also added some additional context, emphasizing that our use
    of travel time estimates to various locations is a standard procedure in
    land use and transport interaction modeling in urban and economic geography,
    where studies commonly model residential choice behavior in a more urban
    context.

6.  Not clear how results of statistical analysis justify conclusion.

    Response: In the results section, we note the following. The binomial GLM is
    a model of ranch locations. The Poisson GLM is a model of ranch size. The
    binomial GLM suggests that ranches are more likely to be occupied the closer
    they are to springs. The exact opposite is true of distance to La Paz. They
    are more likely to be occupied farther from the city. This suggests that
    access to water rather than markets is a bigger constraint on settlement
    locations. But when you look at the size of ranches, the Poisson GLM says
    they are bigger the closer you get to cities and the farther you get from
    springs. So, in the discussion section, we summarize these ideas by noting
    that access to water explains where occupied ranches are located, and access
    to markets explains how big they are.

7.  Increasing wages in the cities may lead rural ranchers to shift time and
    energy away from ranching, decreasing their impacts on local ecology.

    Response: While we don't have data on how much of their income comes from
    cities, we do emphasize the role that demography plays in economics:
    population increases, this leads to increased demand for goods and services,
    and that in turn leads to increased demand for workers to supply those goods
    and services (demand for labor is synonymous with wages). This should
    incentivize movement toward the city, so we should get further levels of
    population decline in these ranching communities. We have added some text to
    emphasize that the consequence of this will be complex and multifaceted.