### 2024-01-24 manuscript/manuscript.qmd
changes to manuscript requested by reviewer

---
title: Trade-off between market and ecosystem services drives settlement decisions among smallholder ranchers in Baja California Sur, Mexico
date: today
author:
  - name: Kenneth B. Vernon
    affiliations: 
      - id: colorado
        address: | 
          Center for Collaborative Synthesis in Archaeology  
          Institute of Behavioral Science  
          University of Colorado, Boulder  
          1440 15th Street  
          Boulder, CO 80302 
  - name: Brian F. Codding
    affiliations:
      - id: utah-anthropology
        address: | 
          University of Utah Department of Anthropology, 
          260 S. Central Campus Drive, RM 4625, 
          Salt Lake City, UT 84112
      - id: gcsc
        address: |
          University of Utah Global Change and Sustainability Center, 
          115 South 1460 East, RM 234 FASB, 
          Salt Lake City, UT 84112
  - name: Simon Brewer
    affiliations:
      - id: utah-geography
        address: |
          University of Utah Department of Geography, 
          260 S. Central Campus Drive, RM 4625, 
          Salt Lake City, UT 84112
  - name: Shane J. Macfarlan
    affiliations:
      - ref: utah-anthropology
      - ref: gcsc
target-journal:
  - "Sustainability Science"
abstract: |
  While smallholder food producers increasingly depend on goods and services provided by distant markets, they are still constrained by their local ecology, as their rural communities often lack many of the infrastructural advantages of urban centers. Navigating the costs and benefits of market and ecosystem services is, as a consequence, crucial to their livelihoods and overall well-being. Here, we explore one aspect of that market-ecology trade-off: its effect on the residential choice behavior of ranchers living in the arid Sierra de la Giganta mountains along the eastern spine of the Baja California peninsula in Mexico. We rely on two Generalized Linear Models (GLMs), a binomial GLM of the locations of occupied and abandoned ranches and a Poisson GLM of population size at occupied ranches. Our proxies for market integration and ecosystem services are travel-time estimates to surficial springs and distant cities. By using these travel time estimates, we enhance the ethnographic study of residential choice behavior in smallholder systems by drawing on key innovations from urban and economic geography. Our models show that ranch clusters are more likely to be abandoned the farther they are from surficial springs, and occupied ranch clusters have larger populations the closer they are to markets. This has important consequences for the ecological resilience of this ranching community, particularly in the context of climate-induced drought, which may lead to ranching strategies that are more locally resilient but less profitable on the market. At the same time, market integration may reshape the social structure of these communities, introducing new incentives that lead them to further integrate with and increasingly depend on those markets. All of these changes are exacerbated by demographic processes, in particular the explosive growth of the region's urban centers.
keywords:
  - Habitat Suitability
  - Land Use 
  - Transportation
  - Sustainability
  - Urbanization
  - Land Abandonment
corresponding_author: Kenneth.Vernon@colorado.edu
linenumbers: false
bibliography: bibliography.bib
format:
  simple-pdf: default
  docx: default
execute: 
  echo: false
  error: false
  warning: false
---

```{r}

library(dplyr)
library(gt)
library(here)
library(knitr)
library(readr)

```

# Introduction

In traditional, small-scale subsistence economies, the value of a habitat is
typically a function of its local resources, particularly their distribution and
abundance [@moran2022]. As individuals integrate with modern markets, however,
that connection becomes more tenuous. This can be a good thing, of course,
allowing individuals to diversify their livelihood, making them more flexible
and capable of absorbing potential risks and uncertainty [@barrett2001;
@little2001]. But, it can have other, unforeseen consequences, too, especially
as individuals come to depend on markets for their livelihood. For those living
in predominantly marginal habitats, it may even have the opposite effect, making
them more sensitive to the effects of climate change and other exogenous forces
like infrastructure development, range fragmentation, and land insecurity
[@fratkin2001; @hobbs2008; @thornton2009]. Together, these conflicting outcomes
suggest important trade-offs for individuals presented with the opportunity to
integrate more closely with modern markets. Time and energy they put into
accessing regional market services is often time and energy they cannot spend
accessing local ecosystem services. How they navigate that trade-off can have
important consequences for their health and well-being [see, for example,
@liebert2013; @urlacher2016; @stagaman2018], including dramatic changes to their
life history and demography [@shenk2016; @colleran2015]. It is, thus, critically
important to understand how and to what degree individuals choose to integrate
with modern markets.

To explore this issue, we focus on a small group of historically marginalized
rural ranchers living in the arid Sierra de la Giganta region of Baja California
Sur (BCS), Mexico [@macfarlan2020a, see @fig-overview-map]. We first explore how
market integration and local ecology structure the distribution of occupied and
abandoned ranch clusters. This is a process that took place over several hundred
years, with the shift to its current distribution beginning in the late 1970s,
following construction of the transpeninsular Federal Highway 1 [@barkenbus1974;
@mathes2015]. We then ask how market integration and local ecology have shaped
the distribution of the population across occupied clusters today.

## Historical Context

While these people are sometimes referred to as "choyeros", a local nickname for
anyone born in the rural, mountainous ranching communities of BCS (also a
reference to the local cholla cactus, *Cylindropuntia cholla*), in this paper,
we refer to them simply as "ranchers" or "rancheros", which are the expressions
they typically use to describe themselves. Descended from Euro-American
colonists who began settling the region in the late-17th century [@crosby1994],
they number approximately 4,000 today, living in communities largely organized
around clusters of one or more ranches housing inter-generational kin
[@macfarlan_etal2020a].

Historically, the isolation of these rural ranching communities in the
mountainous hinterlands of the Sierra de la Giganta meant that getting to the
markets found in nearby urban centers (mostly in the capital of La Paz) was an
extremely costly endeavor, even more so than it is today. This had two important
consequences for these rural ranchers. First, while they do rely on markets for
many of their subsistence needs, local food production still contributes to a
substantial portion of their diet [@macfarlan_etal2023a], meaning they are
neither fully subsistence-based nor fully market integrated. For this reason,
they are more accurately characterized as smallholder ranchers rather than pure
subsistence ranchers or pastoralists [@ifad2013]. Second, their communities
remain some of the most severely under-served in the region, lacking access to
critical services and infrastructure, including piped water, sanitation, and
energy [@secretaríadedesarrollosocial2021].

Given the aridity of the area and lack of regionally connective water
infrastructure, rancheros depend almost entirely on surficial springs to satisfy
their water needs, including drinking water for themselves and their livestock,
as well as irrigation for garden plots (known as huertas and jardines), which
provide a small supplement to human and livestock diets during dry periods
[@macfarlan2020a; @macfarlan_etal2020a; @schniter2021; @lerback2022]. To make
use of springs, ranchers either have to pump water themselves, using gas or
solar powered equipment, or transport it by gravity irrigation [@lerback2022]. A
strong incentive thus exists to locate in closer proximity to these water
resources.

A similar, though competing, incentive exists for accessing markets in urban
areas like Loreto, Ciudad Constitución, and La Paz, the capital of BCS. These
cities furnish rural ranchers with many of the necessities they cannot acquire
locally, including the ability to trade livestock, register animals with the
state, receive medical attention, renew water permits, purchase groceries, and
socialize with people outside their local community. The chance for supplemental
income will also on occasion draw rancheros to the cities, as it has for
pastoralists elsewhere [@barrett2001; @little2001]. This includes working in
construction, seasonal employment in coastal fisheries, or the tourist sector,
specifically within the resort towns of Cabo San Lucas and San Jose del Cabo.

Complicating matters is the fact that land privatization serves as an effective
constraint on these dynamics. This is important because pastoralists have
historically lacked clear title to land, instead relying on open rangeland
systems [@behnke1994; @turner1999]. In Mexico, these systems take the form of
*ejidos*, or areas of communal agricultural land. The first *ejidos* were
established by the Mexican state in the early part of the twentieth century
[@perramond2008], though in our project area they did not arrive until 1975
[@dof1975] and 1976 [@dof1976], respectively. While these *ejidos* have
historically served as a buffer against land fragmentation, allowing rancheros
to maintain their way of life, their ability to play that role continues to
change as the pace of privatization and shifting land tenure accelerates
[@lesorogol2003; @reid_etal2008].

In BCS (along with the rest of Mexico), the Agrarian Land Reforms of 1992 played
an outsized role in driving these changes, as they introduced powerful new
incentives for *ejidos* to parcelize collective land holds and allocate private
rights to individuals [@barnes2009]. While note the sole cause, this new
property regime has increased uncertainty and insecurity about land title and
exposed the region to speculation from real estate markets. As has happened
elsewhere, these privatization efforts appear to be led by NGOs, commercial
interests, and other external forces, with the result being greater
fragmentation of pastoral rangeland systems and increased uncertainty about land
use [@burnsilver_etal2008; @galvin2009]. For the rural ranchers of the Sierra de
la Giganta, this process of fragmentation and increasing insecurity at least
partially explains why ranches of related individuals tend to cluster.
Rancheros, unable to find new land for themselves, simply build ranches next to
their kin on land that is at least historically associated with their family.

To summarize, we want to know how traditional rural ranchers in BCS navigate the
competing costs and benefits of market integration and local ecosystem services.
This is done by estimating their contribution within a regression framework to
the location and size of ranch clusters in two east-west running watersheds
along the southern extent of the Sierra de la Giganta range. Given the extreme
ecological constraints that characterize this region, especially the aridity of
the region, we expect water management to play a substantial role in individual
settlement decisions. Privatization may impose an additional constraint on this
system, especially as it has served to increase land insecurity.

# Methods

## Study Area

As shown in @fig-overview-map, our study area encompasses two east-west running
watersheds with a combined area of roughly 1,042 km$^2$. The area is
approximately equidistant in Cartesian terms between the two nearest urban
centers, being approximately 85 km from Ciudad Constitucion and 95 km from La
Paz. This is an extremely rugged and mountainous terrain (see
@fig-context-photos C), as the elevation of this area ranges from 0 to 980 m,
with a mean of 427 m and a variance of 142 m. The average slope is around 15°.
For reference, just 35 km southwest of the study region, between it and Federal
Highway 1, the average slope over a comparable area is just 0.8°.

As mentioned above, this is also an extremely arid region. It is a southern
extension of the Sonoran Desert Biogeographic Region, so the area is dominated
by columnar cacti and xeric scrub [@shreve_wiggins1964;
@leondelaluz_dominguezcadena2006] capable of sustaining themselves in a place
that averages only 200 mm of rainfall every year [@rebman_roberts2012]. This
circumstance is not helped by the lack of surficial freshwater stocks
[@maya1997], with the only permanent sources of naturally occurring freshwater
being arid-land springs [@grismer1993; @macfarlan2020a; @macfarlan_etal2020a;
@macfarlan2021]. @lerback2022 provides a fine-grained analysis of these spring
systems in the project area, describing them as primarily lentic pools at low
points in the arroyos (see @fig-context-photos B).

## Models

To evaluate the determinants of ranch occupation and population size, we rely on
two generalized linear models (GLMs). We use the first GLM to evaluate whether
market integration or local ecology has played a larger role in structuring the
distribution of occupied and abandoned ranch clusters. For this model we assume
that the dependent variable, whether a ranch cluster is occupied or abandoned,
has a binomial distribution and is related to the linear predictor by the logit
link: $$
\begin{aligned}
Y &\sim B(n, p) \\
E[Y/n] &= p \\
logit(p) &= \beta_{0} + \beta_{1} D(La\; Paz) + \beta_{2} D(Springs)
\end{aligned}
$$ with $n$ being the number of ranch clusters, $p$ the probability that a ranch
cluster is occupied, and $D(x)$ the travel time from a ranch cluster to $x$ (see
below for more details about how we calculate this).

We use the second GLM to identify which variable contributes more to the
distribution of rancher populations across currently occupied clusters, with an
additional variable representing land security, or the proportion of ranch
houses within a cluster having clear title to their property, as this may
mediate ranch population size. For this model, we assume that the dependent
variable, the population size of a ranch cluster, has a negative binomial
distribution and is related to the linear predictor by the log link: $$
\begin{aligned}
Y &\sim NB(\lambda, k) \\
E[Y] &= \lambda \\
log(\lambda) &= \beta_{0} + \beta_{1} D(Cities) + \beta_{2}D(Springs) + \beta_{3} P(Security)
\end{aligned}
$$ with $\lambda$ being the average number of individuals per ranch cluster, $k$
a dispersion parameter, and $P(Security)$ the proportion of ranches in a ranch
cluster that are secure. The use of a negative binomial rather than the standard
Poisson model for count data is preferred here to account for dispersion, or
circumstances in which the variance deviates from the mean, which violates a key
assumption of the Poisson distribution.

Before continuing, we would like to emphasize an important historical dimension
to this research, namely, that the process leading to the current distribution
of occupied and abandoned ranch clusters played out before Ciudad Constitucion
(the large urban area to the north of our study area) offered significant market
benefits. This is reflected in how we specify each model. The model of occupied
and abandoned ranch clusters includes only proximity to La Paz, $D(La\; Paz)$,
as a measure of market integration, which historically was the only market that
ranchers accessed prior to the construction of Federal Highway 1. The model of
population distributions across occupied ranches, however, includes travel time
to La Paz and Ciudad Constitucion, $D(Cities)$, as both have offered markets
regularly used by ranchers in our study population today.

## Data

The dependent variables, occupation status ($p$) and population size
($\lambda$), we obtain from Mexico's online Catálogo do Localidades
[@secretaríadedesarrollosocial2021], coupled with fieldwork conducted over
several years beginning in 2014. In total, the sample includes 72 ranch
clusters, with 39 being unoccupied or abandoned and the remaining 33 being
occupied by 290 individuals. The number of individuals per ranch cluster ranges
from 1 to 32, with a mean of 8.8. To provide additional context,
@fig-context-photos B shows an example of an abandoned ranch.
@fig-context-photos C shows a cluster of currently occupied ranches.

The independent variables, market integration and habitat suitability, we
operationalize using three estimates of travel time: (i) one to La Paz, (ii) one
to Ciudad Constitucion or La Paz (depending on the watershed), and (iii) one to
local springs, with each involving a mix of pedestrian and automotive travel.
The first two variables are our proxies for market integration, under the
assumption that increased interaction with markets will positively covary with
proximity to the cities. The third variable, travel time to springs, is our
proxy for habitat suitability, as water is a critical resource in this arid
environment. In the binomial GLM, which examines the distribution of abandoned
versus active ranching clusters, we use only travel time to La Paz (i) as Ciudad
Constitucion did not offer a significant market when those locations were being
established. We then use travel time to both Ciudad Constitucion and La Paz (ii)
in the negative binomial GLM examining ranch-cluster population sizes. travel
time to springs (iii) is used in both models, on the assumption that the
location of active streams would not have changed during the period of time in
question [@grismer1993].

We define land security as the proportion of ranch houses within a cluster with
clear title to land or who are registered members of an *ejido*. Information on
this is provided by Mexico's National Institute of Statistics and Geography
[@registroagrarionacional2019] and by ethnographic observation. Of the occupied
ranch clusters, only 8 are fully secure, meaning all ranch houses within the
cluster have full title to their property. Of the remaining 25, less than half
of their ranch houses are secure.

A summary of these data can be found in @tbl-variables.

## Travel time

The critical variable in our models is travel time, $D(x)$. This has been used
as a proxy for market integration in other studies focusing on largely
subsistence-based economies [see especially @gurven2017], with models exploring
its relationship to, among other things, social status [@schultz2019], hunting
effort and efficiency [@luz2015], and oral tradition [@schniter2018]. An
important advantage of our analysis, however, is that we use travel time as a
proxy for both market integration and ecosystem services. Specifically, we use
travel time to major cities as a proxy for market integration and travel time to
local surficial springs as a proxy for ecosystem services. This allows for more
straightforward comparisons of their relative contributions to settlement
outcomes, which can then be readily interpreted as strategies for navigating
trade-offs between those two variables. The use of these travel-time estimates
also makes our models analogous to models from urban and economic geography,
where travel time (or, more generally, transport costs) to and from locations
providing various goods and services is assumed to be a key determinant of
residential choice [@alonso1964; @lee_etal2010; @acheampong_silva2015;
@waddell_etal2007].

To estimate travel time, we first derive a gridded cost surface from a 30-meter
resolution digital elevation model obtained from the National Elevation Dataset
provided by the United States Geological Survey. This is achieved by applying
Campbell's hiking function [@campbell2019] to slope estimates extracted from the
elevation model. The result is a cost surface where the value of each 30-meter
grid cell represents the time in seconds it would take to traverse that cell.

Three important assumptions underlie the cost-surface model. First, we assume
that the average walking speed of a rancher is approximately equal to the third
decile of US hikers (about 2.5 meters per second over flat ground), which
increases or decreases based on slope. Second, we assume that no individual will
traverse a slope greater than 35 degrees. Obviously, there will be exceptions to
this, but ethnographic observation does suggest that it is true in general for
the rancheros in this region. Plus, our distance measure is used for relative
comparisons, making the precise absolute value of limited concern. Finally, we
assume that travel on roads and built paths will be less costly than through
grid cells where vegetation and other landform characteristics provide obstacles
to travel. In the cost surface, this is achieved by first classifying paths as
primary (for maintained roads), secondary (for non-maintained roads), and
tertiary (for walking paths), and then weighting the travel cost associated with
each by 0.2, 0.5, and 0.8, respectively. This is equivalent to increasing the
speed of travel to 12.5, 5, and 3.125 m/s, making the speeds more comparable to
driving a car or riding a horse along the primary and secondary paths, which is
what the ranchers here do when going to the city or visiting their springs or
neighbors. All road paths were digitized manually using Google Earth and
validated through field observation.

Once we have this cost surface, we then use Djikstra's algorithm
[@cormen_etal2001] to identify least-cost paths from ranch clusters to springs
and cities and, thus, the total time it takes to travel from each origin to each
destination. These are shown in @fig-shortest-paths. We note, however, that
strictly speaking, we do not measure travel time to the cities, just to the edge
of the study region in the direction of Federal Highway 1, for both the northern
and southern watersheds. Each one has a single road that connects it to Highway
1 and thus to the markets in Ciudad Constitucion to the north and the capital La
Paz to the south. The "road" that connects the watersheds to each other is
poorly maintained, difficult to traverse, and only used out of necessity. As a
consequence, nearly all of the variance in travel time to cities is found within
each watershed. For that reason, we estimate the time it takes to leave each
watershed ("leaving" here means hitting the edge of the project area), rather
than the true time it takes to get to each city. We interpret this as the
relative difference in time it takes to get to Ciudad Constitucion from the
northern watershed and the relative difference in time it takes to get to La Paz
from the southern watershed. To get to La Paz from the northern watershed, we
simply add an arbitrary half hour to the total time to get out of that
watershed. The same would presumably be the case to get to Ciudad Constitucion
from the southern watershed, though that value does not factor into our
analysis. For springs, our ethnographic data reveal that individuals residing
within the same ranch cluster tend to rely on one or two springs at most
[@lerback2022]. Thus, we estimate road distance to the two nearest springs and
take the average.

## Model evaluation

For completeness, we fit three versions of the negative binomial model of ranch
cluster population: one without the land security variable $P(Secure)$, one with
it included as an additive term, and one with it included as an interaction
term. After fitting these models, we perform an ANOVA Likelihood Ratio Test for
significant improvement in fit when adding each version of the land security
variable. We then test for spatial autocorrelation in the untransformed
residuals of all models, including the binomial model, using Monte Carlo
simulations of Moran's I. We also test for multicollinearity in our variables by
calculating their respective variance inflation factors.

All analyses are conducted in the R programming language and environment
[@rcoreteam2023]. For more details on these models and the travel time analysis,
please see the supplementary materials.

# Results

We initiated this research to understand how local ecology and market
integration influenced settlement dynamics. Our first model examines how these
two factors influence whether or not a ranch cluster is occupied or abandoned.
The model suggests that a cluster's status (whether it is occupied or abandoned)
negatively co-varies with travel time to spring ($\beta$ = -4.803, *p* = 0.009)
and positively co-varies with travel time to La Paz ($\beta$ = 2.797, *p* \<
0.001), indicating that ranch clusters are more likely to be occupied closer to
springs and further from town (@tbl-coefficients, @fig-response-plots). Our
second model examines how local ecology and market integration influence the
distribution of the population across occupied ranch clusters today. The model
shows that population size across occupied ranch clusters negatively co-varies
with travel time to cities ($\beta$ = -0.802, *p* \< 0.001), for both
Constitucion and La Paz, but does not significantly co-vary with travel time to
springs ($\beta$ = 0.699, *p* = 0.299) (@tbl-coefficients, @fig-response-plots).
The model with an additive land security term suggests that more secure ranch
clusters have slightly smaller populations, but the effect is non-significant
($\beta$ = -0.114, *p* = 0.687). The ANOVA also indicates that an additive land
security term does not significantly improve the model ($\chi$(1) = 3.003, *p* =
0.684). The same is true for an interactive term ($\chi$(1) = 3.130, *p* =
0.663).

The null hypothesis of no spatial autocorrelation in the untransformed residuals
cannot be rejected for either the occupation status (*I* = -0.023, *p* = 0.820)
or population size (*I* = -0.082, *p* = 0.460) model, thus indicating a lack of
spatial correlation. Measures of variance inflation are minimal in each case,
approximately one for all terms, strongly suggesting no multicollinearity in the
independent variables.

For a complete list of summary statistics for each model, see @tbl-models.

# Discussion

With this analysis, we seek to understand how rural ranchers in BCS respond to
important trade-offs between their local ecology and market integration. Our
results indicate that the locations of modern ranch clusters are largely
constrained by access to surficial water sources, our proxy for habitat
suitability in this arid environment, rather than access to the capital La Paz.
This observation is supported by the fact that occupied ranch clusters are more
likely to be found farther from the capital, rather than closer to it. However,
within those constraints, the rancheros are now living at larger densities in
locations with easier access to Ciudad Constitucion and La Paz, our proxies for
market integration. In other words, water availability determines *where* people
live but not *how many* people live there. Instead, that is being driven largely
by markets, which pull people closer to the cities.

## Proxies for market integration

One important advantage of this research concerns the use of travel time as a
measure of both market integration and local adaptation. Using this simple
metric has a number of advantages. First, it provides a straightforward way of
evaluating the relative contribution of each variable to the outcome of
interest, in this case, settlement decisions among rural ranchers. Second, it
conforms with geographic theory regarding spatial interactions and dependence
(otherwise known as Tobler's Law, @tobler1970) and overcomes known limitations
of Euclidean distance as a cost measure. Finally, it is a relatively simple
matter to derive those cost estimates from easily accessible remote-sensed
elevation and road network data held in online repositories. High resolution
informant data, on the other hand, are generally difficult and costly to
acquire, often with little chance of providing the right sort of overlap between
different variables required for fair comparisons.

It is true, as @mattison2022 point out, that market integration is a
multi-dimensional process, making it at first blush an awkward concept to try
and capture in a single variable like travel time or caloric-intake
[@henrich2010]. Still, Tobler's Law would seem to suggest that all the variables
strongly associated with market integration will also co-vary with proximity to
cities and their markets. In other words, the costs and benefits of market
integration will positively co-vary with access to markets - both will increase
as one gains greater access to and interacts more often with markets. So, the
closer individuals are, the more likely they should be to derive their caloric
requirements from cities, to work in cities, to own televisions, cellphones, and
other goods sold in cities, but also the more likely they should be to face
intense competition for jobs, higher rents, and other costs associated with
living in that proximity.

The ranchers in our study population also recognize that those who live closer
to the trans-peninsular highway have greater access to markets and are more
likely to interact with them. They have said as much during our conversations
with them. As mentioned above, travel time is also a fundamental variable in
urban and economic geography [@krugman1991; @fujita2001] and thus finds support
from that direction, too. So, when we look at where individuals actually choose
to live (how near or how far from cities and markets they choose to reside), it
is safe to assume we are also getting an answer to the question of how they
weigh the costs and benefits of cities against those afforded by a rural way of
life.

## Historical antecedents

What explains our results? There are many dimensions to this question, but key
to any general understanding of such processes is the idea that people will
choose to live where they can make the best living for themselves given the
socioeconomic and environmental constraints that confront them. Those
constraints will, in turn, differ depending on how one chooses to go about
making a living [@vernon2022]. For the people of the Sierra de la Giganta, that
means living where you can expect to do ranching and basic crop production
reasonably well, that is, where one has access to permanent water sources, as
well as flat lands that sit above dry riverbeds (so they are not impacted by
flooding caused by seasonal tropical storms and hurricanes). Due to the lack of
meaningful water infrastructure in this region, households depend on these
springs to support both domestic (e.g. bathing and drinking) and economic needs
(e.g. supporting livestock and crop production). Our statistical results are
consistent with this idea and with our ethnographic interviews with ranchers who
indicate that access to a spring is the single most important factor for
establishing and maintaining a home [@lerback2022].

Our first statistical model indicates that abandoned ranches were located
farther from springs. Although this appears to run counter to our previous claim
about the importance of springs to all settlement in the region, we believe this
is an artifact of our analytic framework, which required us to collapse the
temporal dimension of this largely historical process. Historical and
ethnographic evidence suggest that as the first Euro-American colonists entered
this region in the 18th century, they occupied the largest springs in close
proximity to arable land [@crosby1994; @macfarlan_etal2020;
@macfarlan_etal2021]. As the population expanded over time, individuals would
move up or down drainages, to the next closest spring that could support
subsistence needs. After time, ranchers likely saturated the landscape by
occupying all areas with permanent freshwater [sensu @crosby1981] as well as
less desirable locations which were supported not by springs, but by ephemeral
freshwater sources such as tinajas and seasonally filled lake basins, which can
hold water for months or years if they receive sufficient precipitation via the
North American Monsoon.

However, after the construction of the Transpeninsular Highway and associated
graded dirt roads in the 1970s and 1980s, the payoffs structuring settlement
decisions were likely altered, as novel constraints and opportunities emerged
[@barkenbus1974; @mathes2015]. For a smallholder population undergoing market
transition, the ideal household location has become both next to a spring and
close to a road. These roads allowed ranchers to access urban goods and services
more easily and created an incentive for households occupying springs and
ephemeral freshwater sources far from the road to move closer to it
[@macfarlan_etal2023a]. This process resulted in a depopulation event between
the two major drainages. As this process has unfolded over the last 30-40 years,
springs near roads with the closest access to urban markets have experienced
increasing intensification as suitable habitats for establishing new households
have been exhausted, resulting in the highest population densities.

## Land Security

Our measure of land security proved to be non-significant in our model. Rather
than being an outcome from which to draw conclusions, we believe this result
should be interpreted with caution. Our measure of land security is based on a
combination of government data and informant observations and so likely does not
capture the full complexity of this important issue. A larger dataset with a
more standardized and quantifiable measure of land security would be more
representative. We also note that the issue of land security is not equally
salient to everyone in this population. Many of the informants that might be
land insecure by most objective measures do not necessarily act as if their
title to land is uncertain. This would suggest that perceptions of risk with
respect to property rights should be studied more thoroughly.

## Socio-ecological consequences

The findings presented here have important consequences for traditional ranching
communities living in similar circumstances, for how they maintain their way of
life while meeting the challenges posed by climate change, land fragmentation,
and other exogenous forces. For instance, researchers have shown that market
integration within traditional agricultural communities involves a shift from
intensive to extensive kinship systems, or shifts from small kin networks with
higher degrees of relatedness to large kin networks with lower degrees of
relatedness [@peña2002]. This appears to be the result of a decreasing reliance
on consanguineous or close-kin marriage customs, with families favoring
expansive social networks within a market economy over stronger family ties
within a largely agrarian economy [@shenk2016]. Currently, the ranching
communities of BCS are characterized by more intensive kin networks, but the
limited available evidence suggests that this might be changing, with
individuals increasingly marrying outside of the ranching community, most likely
to individuals living within the cities. Families are as a consequence
increasing their connections to the cities and by extension increasing the
incentives for integrating more closely with their markets.

These settlement dynamics have also had variable ecosystem outcomes. Springs
that exist closer to the road, which are more heavily utilized, typically
experience substantial negative outcomes relative to springs that exist in the
depopulated zone, including biodiversity loss, higher rates of introduced exotic
species, soil degradation, and noise and chemical pollution
[@leondelaluz_dominguezcadena2006]. This is unfortunate, as the springs and the
emergent terrestrial ecosystems they support are considered Ramsar sites
(wetlands of international importance), as well as national terrestrial and
hydrological priority regions [@arriagacabrera2002; @ramsar2022]. However, it is
inaccurate to assume, as many conservation biologists, NGOs, and urban elites
seem to do, that ranchers and their lifestyles have uniformly negative impacts
on regional ecosystem health [@riemann2005]. For it appears that small, riparian
ecosystems surrounding springs in the depopulated hinterland away from roads are
currently on the rebound, with higher rates of endemism, and this in spite of
the fact that ranching continues in the region. Similar outcomes have also been
noted elsewhere [@vaccaro2021]. So, concentration of the population within this
region might actually be beneficial, at least from the perspective of
biodiversity and ecosystem health, a pattern that appears to be recurring across
the globe as populations become more urbanized [@daskalova2023]. For that
matter, current research suggests that the complete abandonment of a region can
have negative consequences for biodiversity given the co-evolutionary histories
of small-scale subsistence economies and their local ecosystems [@fischer2012].

Current demographic trends across the region may also exacerbate these local
processes [@vasco2017]. Since the signing of NAFTA, the total population of the
southern peninsula has more than doubled, from 375 thousand in 1995 to nearly
800 thousand in 2020. As one might expect, these increases have occurred almost
exclusively within the peninsula's urban corridors [@inegi2023a; @inegi2023] .
This entails increased demand for goods and services and by necessity, increased
demand for the labor that can provide them, thereby increasing the incentives
for locating in closer proximity to cities and abandoning more remote locations.
As alluded to above, the ecological consequences will be complex and
multifaceted. Given the complexities of population decline and its consequences
for biodiversity, it is hard to say whether these demographic trends will have a
net positive or negative effect on long term ecosystem health. We hope our
analyses here catalyze future research on the effects of ranch abandonment on
ecosystem outcomes.

## Climate change and drought

Of particular urgency for the rural ranchers of the Sierra de la Giganta is
ongoing climate-induced drought [@cavazos2012; @colorado-ruiz2018]. For the
reasons we have emphasized throughout, a lack of access to government services
and infrastructure makes them uniquely sensitive to these local impacts
[@hansen2019]. That is almost certainly why access to water plays a larger role
in constraining the location of occupied ranch clusters. It may also explain the
decisions ranchers make about the size and composition of their herds, as intra-
and inter-annual variation in precipitation - in the absence of water
infrastructure - is the primary factor driving the abundance and distribution of
most animal fodder in the region [@pettorelli2013; @macfarlan_etal2023a].
Previous research has shown that pastoralists respond to and oftentimes prepare
for drought conditions by reducing their herd size, selling off stock in the
nearby cities, or switching to more drought tolerant alternatives like small
ruminants [@mace1989; @mace1990; @mace1993]. While these animals are more
efficient in terms of their total land area, feed, and water requirements, there
is less demand for them in city markets. This implies that ranchers in the
region, while trying to address the ongoing threat of droughts, will lose
important sources of capital they would need to access other resources provided
by modern markets that they depend on.

## Conclusion

Trade-offs between local ecology and market integration will continue to shape
the development of rural, smallholder communities, especially as globalization
and urbanization accelerate. This is as true of the rural ranchers of the Sierra
de la Giganta as it is for ranchers, pastoralists, and small-scale
agriculturalists in other parts of the world. For ranchers living in the remote
corners of BCS, the data coupled with informant reporting together suggest that
these ranchers feel the pull of markets, being compelled to infill around
ranches nearer to roads that give them greater access to those markets. The
consequences of this spatial sorting of the population are complicated as well.
While integration can help to mitigate local risks, it also raises new ones,
with population packing increasing pressure on nearby water resources that are
crucial to the sustainability of these communities. At the same time, the
concentration of the population appears to offer positive results for the
ecology of abandoned areas, among other things increasing their biodiversity.

The research presented in this paper raises a number of important issues that
deserve greater attention, including a more robust analysis of land
fragmentation and land insecurity, using a larger dataset more representative of
these ranching communities across the peninsula. It is also important that we
better understand the manner and extent to which proximity or travel time to
markets correlates with other proxies for market integration. Answering that
question would go along way towards helping researchers overcome many of the
methodological challenges associated with richly ethnographic studies of remote,
smallholder economies and in the process offer new avenues for investigating the
ongoing integration of those systems with modern markets.

{{< pagebreak >}}

# Funding and/or Competing interests

The research was funded in part by the Center for Latin American Studies,
University of Utah.

The authors have no relevant financial or non-financial interests to disclose.

{{< pagebreak >}}

# References

::: {#refs}
:::

{{< pagebreak >}}

```{r}
#| label: fig-overview-map
#| fig-align: center
#| out-width: 5.75in
#| fig-cap: "**Overview map.** The top right panel shows the location of the study area (in red) in relation to the two nearest city centers, La Paz and Ciudad Constitucion. Panel A shows the distribution of ranch clusters in each watershed, as well as their population size (size and color). Note that a population of zero means the ranch cluster is abandoned. Panel B shows the distribution of land secure and land insecure ranch houses (shape and color) for all occupied ranch houses. Solid white lines represent Federal Highway 1 (in the top right panel), major roads connecting it to the project area, and other roads and paths in the project area."

here("figures", "overview-map.png") |> include_graphics()

```

{{< pagebreak >}}

```{r}
#| label: fig-shortest-paths 
#| fig-align: center 
#| out-width: 5.5in 
#| fig-cap: "**Shortest paths.** Maps showing least-cost paths to springs and cities."

here("figures", "least-cost-paths.png") |> include_graphics()

```

{{< pagebreak >}}

```{r}
#| label: fig-response-plots
#| fig-align: center
#| out-width: 5.5in
#| fig-cap: "**Response plots.** Marginal response plots showing change in the dependent variables (occupied vs abandoned ranch clusters and population size) as a function of change in one independent variable while holding the other at its mean."

here("figures", "response-plots.png") |> include_graphics()

```

{{< pagebreak >}}

```{r}
#| label: fig-context-photos
#| fig-align: center
#| out-width: 5.5in
#| fig-cap: "**Context photographs.** (A) Spring San Francisco, an oasis around a lentic pool in a shallow arroyo. (B) Abandoned ranch near Spring San Francisco. (C) Aerial overview of Arroyo La Presa and a cluster of ranches in the western part of the study. The area is criss-crossed by graded and two-track dirt roads, with a lentic pool in the foreground and huertas or small gardens in the background."

here("figures", "context-photos.png") |> include_graphics()

```

{{< pagebreak >}}

```{r}

table_defaults <- function(x){

  x |> 
    tab_options(
      column_labels.font.weight = "bold",
      data_row.padding = px(5),
      heading.align = "left",
      heading.border.bottom.style = "hidden",
      row_group.as_column = TRUE,
      table.border.top.style = "hidden",
      table.border.bottom.style = "hidden"
    ) |> 
    sub_missing(missing_text = "")

}

```

```{r}
#| label: tbl-variables
#| fig-align: center
#| out-width: "100%"

fn <- here("_misc", "tbl-variables.png")

here("data", "variable-summary.csv") |> 
  read_csv() |> 
  mutate(
    description = case_when(
      variable == "paz"        ~ "Time to leave community on way to La Paz (hours)",
      variable == "cities"     ~ "Time to leave community on way to La Paz from south 
                                  or Ciudad Constitucion from north (hours)",
      variable == "springs"    ~ "Mean time to two nearest springs (hours)",
      variable == "secure"     ~ "Proportion of cluster with clear title or *ejido* 
                                  membership (0 to 1)",
      variable == "occupied"   ~ "Cluster has at least one resident (1) or no 
                                  residents (0)",
      variable == "population" ~ "Number of individuals residing at ranch cluster"
    )
  ) |> 
  select(model, variable, description, everything()) |> 
  group_by(model) |> 
  gt() |> 
  table_defaults() |>  
  tab_stubhead("model") |>
  tab_header(title = "Variable Summary") |> 
  fmt_number(decimals = 2) |> 
  fmt_markdown(description) |> 
  cols_align("left", description) |> 
  gtsave(filename = fn)

fn |> include_graphics()

```

{{< pagebreak >}}

```{r}
#| label: tbl-models
#| fig-align: center
#| out-width: "80%"

here("data", "model-statistics.rdata") |> load()

fn <- here("_misc", "tbl-models.png")

model_statistics |>
  gt() |> 
  table_defaults() |> 
  tab_header(title = "Summary Statistics") |> 
  gtsave(filename = fn)

fn |> include_graphics()

```

{{< pagebreak >}}

```{r}
#| label: tbl-coefficients
#| fig-align: center
#| out-width: "65%"

fn <- here("_misc", "tbl-coefficients.png")

coefficient_estimates |>
  group_by(model) |> 
  gt() |> 
  table_defaults() |> 
  tab_stubhead("model") |> 
  tab_header(title = "Coefficient Estimates") |> 
  gtsave(filename = fn)

fn |> include_graphics()

```